---
title: "Comparing PMcC's `sparsemle` and Matthew's `ashr`"
author: "Lei Sun"
date: "2019-01-16"
output:
  workflowr::wflow_html:
    code_folding: hide
---

## Introduction

This is a quick simulation to compare PMcC's `sparsemle` and Matthew's `ashr`. The observations are signals plus Gaussian noise, and the signals are 90\% zero.
$$
\begin{aligned}
z_j &\overset{\text{iid}}{\sim} N(\theta_j, 1)\\
\theta_j &\overset{\text{iid}}{\sim} g = 0.9\delta_0 + 0.1 g_1
\end{aligned}
$$

The quantities we are comparing for each $j = 1, \ldots, p$ are `lfpr` from `sparsemle` which is computed as
$$
\begin{aligned}
\text{lfpr}(z_j)=\frac{1-\rho}{1-\rho+\rho\zeta(z_j, d)},
\end{aligned}
$$
and `lfdr` from `ashr` which is
$$
\begin{aligned}
\text{lfdr}(z_j)=\text{Pr}(\theta_j = 0|z_j, \hat g)
\end{aligned}
$$

```{r, fig.show = 'hide'}
source("http://www.stat.uchicago.edu/~pmcc/courses/sparsity.R")
library(ashr)
```

## Unimodal signals

Here $g_1 = N(0, 3^2)$.

```{r}
set.seed(777)
theta <- sample(c(rep(0, 900), rnorm(100, 0, 3)))
z <- theta + rnorm(1000)
fit.sparsemle <- sparsemle(z)
fit.ashr <- ashr::ash(z, 1)
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
hist(theta, xlab = expression(theta))
plot(lfpr, inactivity_rate(fit.sparsemle), xlab = "lfpr (Sparse MLE)", ylab = "inactivity rate (Sparse MLE)", main = "lfpr vs inactivity rate")
abline(0, 1, col = "blue", lty = 2)
plot(ashr::get_lfdr(fit.ashr), ashr::get_lfsr(fit.ashr), xlab = "lfdr (ashr)", ylab = "lfsr (ashr)", main = "lfdr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfdr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfdr (ashr)", main = "lfpr vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfsr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfsr (ashr)", main = "lfpr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfdr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfdr (ashr)", main = "inactivity rate vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfsr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfsr (ashr)", main = "inactivity rate vs lfsr")
abline(0, 1, col = "blue", lty = 2)
```

## Point-mass signals

Here $g_1 = \delta_3$, a point mass at $3$.

```{r}
set.seed(777)
theta <- sample(c(rep(0, 900), rep(3, 100)))
z <- theta + rnorm(1000)
fit.sparsemle <- sparsemle(z)
fit.ashr <- ashr::ash(z, 1)
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
hist(theta, xlab = expression(theta))
plot(lfpr, inactivity_rate(fit.sparsemle), xlab = "lfpr (Sparse MLE)", ylab = "inactivity rate (Sparse MLE)", main = "lfpr vs inactivity rate")
abline(0, 1, col = "blue", lty = 2)
plot(ashr::get_lfdr(fit.ashr), ashr::get_lfsr(fit.ashr), xlab = "lfdr (ashr)", ylab = "lfsr (ashr)", main = "lfdr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfdr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfdr (ashr)", main = "lfpr vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfsr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfsr (ashr)", main = "lfpr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfdr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfdr (ashr)", main = "inactivity rate vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfsr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfsr (ashr)", main = "inactivity rate vs lfsr")
abline(0, 1, col = "blue", lty = 2)
```

## No exact null

We now make a small change to the setting: here $g = 0.9N(0, 0.5^2) + 0.1g_1$, i.e., no $\theta_j$ that is exactly zero.

### Unimodal signals

$g_1 = N(0, 3^2)$

```{r}
set.seed(777)
theta <- sample(c(rnorm(900, 0, 0.5), rnorm(100, 0, 3)))
z <- theta + rnorm(1000)
fit.sparsemle <- sparsemle(z)
fit.ashr <- ashr::ash(z, 1)
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
hist(theta, xlab = expression(theta))
plot(lfpr, inactivity_rate(fit.sparsemle), xlab = "lfpr (Sparse MLE)", ylab = "inactivity rate (Sparse MLE)", main = "lfpr vs inactivity rate")
abline(0, 1, col = "blue", lty = 2)
plot(ashr::get_lfdr(fit.ashr), ashr::get_lfsr(fit.ashr), xlab = "lfdr (ashr)", ylab = "lfsr (ashr)", main = "lfdr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfdr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfdr (ashr)", main = "lfpr vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfsr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfsr (ashr)", main = "lfpr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfdr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfdr (ashr)", main = "inactivity rate vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfsr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfsr (ashr)", main = "inactivity rate vs lfsr")
abline(0, 1, col = "blue", lty = 2)
```

### Point-mass signals

```{r}
set.seed(777)
theta <- sample(c(rnorm(900, 0, 0.5), rep(3, 100)))
z <- theta + rnorm(1000)
fit.sparsemle <- sparsemle(z)
fit.ashr <- ashr::ash(z, 1)
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
hist(theta, xlab = expression(theta))
plot(lfpr, inactivity_rate(fit.sparsemle), xlab = "lfpr (Sparse MLE)", ylab = "inactivity rate (Sparse MLE)", main = "lfpr vs inactivity rate")
abline(0, 1, col = "blue", lty = 2)
plot(ashr::get_lfdr(fit.ashr), ashr::get_lfsr(fit.ashr), xlab = "lfdr (ashr)", ylab = "lfsr (ashr)", main = "lfdr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfdr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfdr (ashr)", main = "lfpr vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfsr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfsr (ashr)", main = "lfpr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfdr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfdr (ashr)", main = "inactivity rate vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfsr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfsr (ashr)", main = "inactivity rate vs lfsr")
abline(0, 1, col = "blue", lty = 2)
```

## Fat-tailed signals

As suggested by PMcC, we now try this more sparse and more fat-tailed signal distribution:
$g = 0.95\delta_0 + 0.05 t_2$

```{r}
set.seed(777)
theta <- sample(c(rep(0, 950), rt(50, 2)))
z <- theta + rnorm(1000)
fit.sparsemle <- sparsemle(z)
fit.ashr <- ashr::ash(z, 1)
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
hist(theta, xlab = expression(theta))
plot(lfpr, inactivity_rate(fit.sparsemle), xlab = "lfpr (Sparse MLE)", ylab = "inactivity rate (Sparse MLE)", main = "lfpr vs inactivity rate")
abline(0, 1, col = "blue", lty = 2)
plot(ashr::get_lfdr(fit.ashr), ashr::get_lfsr(fit.ashr), xlab = "lfdr (ashr)", ylab = "lfsr (ashr)", main = "lfdr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfdr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfdr (ashr)", main = "lfpr vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(lfpr, ashr::get_lfsr(fit.ashr), xlab = "lfpr (Sparse MLE)", ylab = "lfsr (ashr)", main = "lfpr vs lfsr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfdr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfdr (ashr)", main = "inactivity rate vs lfdr")
abline(0, 1, col = "blue", lty = 2)
plot(inactivity_rate(fit.sparsemle), ashr::get_lfsr(fit.ashr), xlab = "inactivity rate (Sparse MLE)", ylab = "lfsr (ashr)", main = "inactivity rate vs lfsr")
abline(0, 1, col = "blue", lty = 2)
```

## Comparing `lfsr` (`ashr`) and `lfpr` (`sparsemle`)

In order to generate visible difference between the estimated `lfdr` and `lfsr` from `ashr`, in this simulation $\theta_j \sim g = N(0, 2^2)$, and so $g$ has no point mass at zero. We first compute `lfpr` from `sparsemle`.

When applying `ashr` we are using the following two settings:

1. `method = "shrink"`: when estimating $g$, `ashr` doesn't impose a point mass at zero. Under this setting, `lfdr` will be 0 for all $j$, but we can compute `lfsr`, which is defined as
$$
\text{lfsr}_j := \min\{
\text{Pr}(\theta_j \geq 0 \mid Z_j, \hat g), \ 
\text{Pr}(\theta_j \leq 0 \mid Z_j, \hat g)
\}
$$
2. `method = "fdr"`: when estimating $g$, `ashr` does impose a point mass at zero and encourage conservative over-estimation of this point mass proportion. Under this setting, we can compute both `lfdr` and `lfsr`. Presumbly, these two estimates should be visibly different.

In addition, note that under the Bayesian oracle $\theta \sim N(0, 2^2), Z \sim N(\theta, 1)$, `lfsr` can be computed analytically as $\Phi(-2|Z|/\sqrt5)$.

To sum up, in this simulation we are comparing five quantities:

    - the oracle `lfsr`
    - the estimated `lfsr` from `ashr` using non point mass
    - the estimated `lfsr` from `ashr` using point mass
    - the estimated `lfdr` from `ashr` using point mass
    - the `lfpr` from `sparsemle`.

We put the oracle `lfsr` on the x axis, and other four quantities on the y axis to compare them all.

```{r}
set.seed(777)
theta <- rnorm(1000, 0, 2)
z <- theta + rnorm(1000)

hist(theta, xlab = expression(theta))

fit.sparsemle <- sparsemle(z)
fit.ashr.shrink <- ashr::ash(z, 1, method = "shrink")
fit.ashr.fdr <- ashr::ash(z, 1, method = "fdr")
lfsr.oracle <- pnorm(-2 * abs(z) / sqrt(1 + 2^2))
lfpr <- (1 - fit.sparsemle$rho) / (1 - fit.sparsemle$rho + fit.sparsemle$rho * zeta_ip(z))
lfsr.shrink <- ashr::get_lfsr(fit.ashr.shrink)
lfsr.fdr <- ashr::get_lfsr(fit.ashr.fdr)
lfdr.fdr <- ashr::get_lfdr(fit.ashr.fdr)
order.lfsr.oracle <- order(lfsr.oracle)

plot(lfsr.oracle, lfsr.shrink, pch = 19, cex = 0.5, xlim = c(0, 0.5), ylim = c(0, 0.5),
     xlab = "lfsr (Oracle)", ylab = "other quantities")
points(lfsr.oracle, lfsr.fdr, pch = 19, cex = 0.5, col = 4)
points(lfsr.oracle, lfdr.fdr, pch = 19, cex = 0.5, col = 3)
points(lfsr.oracle, lfpr, pch = 19, cex = 0.5, col = 2)
legend("bottomright", col = c(1, 4, 3, 2), pch = 19, c(
  "lfsr (no point mass)",
  "lfsr (point mass)",
  "lfdr (point mass)",
  "lfpr (sparsemle)"
))
```

It appears that

1. The `lfsr` from `ashr` using no point mass at zero (black) are very close to the oracle, which shows that `ashr` as an empirical Bayes method works pretty well when $g$ is correctly specified.
2. All other three quantities are more conservative than the oracle.
3. `lfpr` from `sparsemle` appears to be most closely related to `lfsr` from `ashr` using (a misspecified) point mass at zero.

## Remarks

`sparsemle` provides `lfpr` and `inactivity rate`, while `ashr` provides `lfdr` and `lfsr`. All four numbers are quantifying the extent to which an observation $Z_j$ is associated with a non-signal. Generally speaking, `inactivity rate` is more conservative than `lfpr`, which is more conservative (when it is small) than `lfdr` and `lfsr`. The difference between `inactivity rate` and `lfpr` could be notable in some simulation settings, while in all of our settings, `lfpr` is quite comparable to `lfdr` and `lfsr`, and `lfdr` and `lfsr` are essentially indistinguishable. Interestingly, `lfpr` and `lfdr` (`lfsr`) are most similar when we have unimodal signals and no exact nulls.
